Distributional Trees and Forests

### 2017-10-10 pkg/disttree/inst/draft/draft.Rnw
variable importance: barplot of top 10 instead of plot of marginal effects on distribution parameter
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1649 255f4249-2263-4417-b1ea-8498779e27a7

@@

  Reto Stauffer \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"at Innsbruck \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Reto.Stauffer@uibk.ac.at} \\
  URL: \url{https://www.uibk.ac.at/statistics/personal/stauffer/reto-stauffer.html}\\
}

@@

Analyzing variable importance of the distforest model leads to the following list of the covariates which show the highest values in variable importance.\\
In the following plot the marginal effect of these variables on the distribution parameters are plotted. (All other covariates are kept at their mean.)
<<variabble_importance, echo=FALSE, eval=TRUE>>=
if(file.exists("vimp100_crps.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/vimp100_crps.rda")
  load("vimp100_crps.rda")
} else {

@@

  #vimp_crps <- varimp(res$df, nperm = 1L)
  #vimp10_crps <- varimp(res$df, nperm = 10L)
  vimp100_crps <- varimp(res$df, nperm = 100L)
  #vimp1000_crps <- varimp(res$df, nperm = 1000L)

  #save(vimp_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp_crps.rda")
  #save(vimp10_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp10_crps.rda")
  save(vimp100_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp100_crps.rda")
  #save(vimp1000_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp1000_crps.rda")

@@

most_impvar <- sort(vimp100_crps, decreasing = TRUE)[c(1:20)]
most_impvar <- most_impvar[most_impvar>2.7]
names_mi <- names(most_impvar)
covnames <- colnames(res$learndata)
id_mi <- match(names_mi, covnames)

datafix <- res$learndata
datafix1 <- datafix2 <- datafix3 <- datafix4 <- 
  datafix5 <- datafix5 <- datafix6 <- datafix7 <- datafix8 <- datafix

# fix all variables except for the most important ones to their mean
for(i in 3:NCOL(datafix))  if(i!=id_mi[1]) datafix1[,i] <- rep(mean(datafix[,i], na.rm = TRUE), NROW(datafix))
for(i in 3:NCOL(datafix))  if(i!=id_mi[2]) datafix2[,i] <- rep(mean(datafix[,i], na.rm = TRUE), NROW(datafix))
for(i in 3:NCOL(datafix))  if(i!=id_mi[3]) datafix3[,i] <- rep(mean(datafix[,i], na.rm = TRUE), NROW(datafix))
for(i in 3:NCOL(datafix))  if(i!=id_mi[4]) datafix4[,i] <- rep(mean(datafix[,i], na.rm = TRUE), NROW(datafix))
for(i in 3:NCOL(datafix))  if(i!=id_mi[5]) datafix5[,i] <- rep(mean(datafix[,i], na.rm = TRUE), NROW(datafix))
for(i in 3:NCOL(datafix))  if(i!=id_mi[6]) datafix6[,i] <- rep(mean(datafix[,i], na.rm = TRUE), NROW(datafix))
for(i in 3:NCOL(datafix))  if(i!=id_mi[7]) datafix7[,i] <- rep(mean(datafix[,i], na.rm = TRUE), NROW(datafix))
for(i in 3:NCOL(datafix))  if(i!=id_mi[8]) datafix8[,i] <- rep(mean(datafix[,i], na.rm = TRUE), NROW(datafix))

pred1 <- predict(res$df, newdata  = datafix1, type = "parameter")
plotdata1 <- cbind(datafix1, pred1)
sp1 <- plotdata1[order(datafix1[,id_mi[1]]),]

# plot effect of most important variables on parameter
if(FALSE){

  pred1 <- predict(res$df, newdata  = datafix1, type = "parameter")
  plotdata1 <- cbind(datafix1, pred1)
  sp1 <- plotdata1[order(datafix1[,id_mi[1]]),]
  plot(x = sp1[,id_mi[1]], y = sp1[, "mu"], type = "l", xlab = names_mi[1], ylab = "mu")
  plot(x = sp1[,id_mi[1]], y = sp1[, "sigma"], type = "l", xlab = names_mi[1], ylab = "sigma")

  pred2 <- predict(res$df, newdata  = datafix2, type = "parameter")
  plotdata2 <- cbind(datafix2, pred2)
  sp2 <- plotdata2[order(datafix2[,id_mi[2]]),]
  plot(x = sp2[,id_mi[2]], y = sp2[, "mu"], type = "l", xlab = names_mi[2], ylab = "mu")
  plot(x = sp2[,id_mi[2]], y = sp2[, "sigma"], type = "l", xlab = names_mi[2], ylab = "sigma")

  pred3 <- predict(res$df, newdata  = datafix3, type = "parameter")
  plotdata3 <- cbind(datafix3, pred3)
  sp3 <- plotdata3[order(datafix3[,id_mi[3]]),]
  plot(x = sp3[,id_mi[3]], y = sp3[, "mu"], type = "l", xlab = names_mi[3], ylab = "mu")
  plot(x = sp3[,id_mi[3]], y = sp3[, "sigma"], type = "l", xlab = names_mi[3], ylab = "sigma")

  pred4 <- predict(res$df, newdata  = datafix4, type = "parameter")
  plotdata4 <- cbind(datafix4, pred4)
  sp4 <- plotdata4[order(datafix4[,id_mi[4]]),]
  plot(x = sp4[,id_mi[4]], y = sp4[, "mu"], type = "l", xlab = names_mi[4], ylab = "mu")
  plot(x = sp4[,id_mi[4]], y = sp4[, "sigma"], type = "l", xlab = names_mi[4], ylab = "sigma")

  pred5 <- predict(res$df, newdata  = datafix5, type = "parameter")
  plotdata5 <- cbind(datafix5, pred5)
  sp5 <- plotdata5[order(datafix5[,id_mi[5]]),]
  plot(x = sp5[,id_mi[5]], y = sp5[, "mu"], type = "l", xlab = names_mi[5], ylab = "mu")
  plot(x = sp5[,id_mi[5]], y = sp5[, "sigma"], type = "l", xlab = names_mi[5], ylab = "sigma")

  pred6 <- predict(res$df, newdata  = datafix6, type = "parameter")
  plotdata6 <- cbind(datafix6, pred6)
  sp6 <- plotdata6[order(datafix6[,id_mi[6]]),]
  plot(x = sp6[,id_mi[6]], y = sp6[, "mu"], type = "l", xlab = names_mi[6], ylab = "mu")
  plot(x = sp6[,id_mi[6]], y = sp6[, "sigma"], type = "l", xlab = names_mi[6], ylab = "sigma")

  pred7 <- predict(res$df, newdata  = datafix7, type = "parameter")
  plotdata7 <- cbind(datafix7, pred7)
  sp7 <- plotdata7[order(datafix7[,id_mi[7]]),]
  plot(x = sp7[,id_mi[7]], y = sp7[, "mu"], type = "l", xlab = names_mi[7], ylab = "mu")
  plot(x = sp7[,id_mi[7]], y = sp7[, "sigma"], type = "l", xlab = names_mi[7], ylab = "sigma")

  pred8 <- predict(res$df, newdata  = datafix8, type = "parameter")
  plotdata8 <- cbind(datafix8, pred8)
  sp8 <- plotdata8[order(datafix8[,id_mi[8]]),]
  plot(x = sp8[,id_mi[8]], y = sp8[, "mu"], type = "l", xlab = names_mi[8], ylab = "mu")
  plot(x = sp8[,id_mi[8]], y = sp8[, "sigma"], type = "l", xlab = names_mi[8], ylab = "sigma")

@@

\begin{figure}[t!]
\begin{subfigure}{0.5\textwidth}
<<rain_axams_varim_margeffects1_mu, fig=TRUE, echo=FALSE>>=
plot(x = sp1[,id_mi[1]], y = sp1[, "mu"], type = "l", xlab = names_mi[1], ylab = "mu")
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.5\textwidth}
<<rain_axams_varim_margeffects1_sigma, fig=TRUE, echo=FALSE>>=
plot(x = sp1[,id_mi[1]], y = sp1[, "sigma"], type = "l", xlab = names_mi[1], ylab = "sigma")
@
\end{subfigure}
\caption{\label{varimp}marginal effect of covariate with highest variable importance in distforest}
\end{figure}
(other covariates? no effects for 4th variable, probably because only splits after several other splits)

>>>>

  Reto Stauffer \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"at Innsbruck \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Reto.Stauffer@uibk.ac.at} \\
  URL: \url{https://www.uibk.ac.at/statistics/personal/stauffer/reto-stauffer.html}\\

@@

@@

Analyzing variable importance of the distforest model leads to the following list of the covariates which show the highest values in variable importance.\\
In the following plot the marginal effect of these variables on the distribution parameters are plotted. (All other covariates are kept at their mean.)
<<variabble_importance, echo=FALSE, eval=TRUE>>=
if(file.exists("vimp_crps.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/vimp_crps.rda")
  load("vimp_crps.rda")
} else {

@@

  vimp_crps <- varimp(res$df, nperm = 1L)
  #vimp10_crps <- varimp(res$df, nperm = 10L)
  #vimp100_crps <- varimp(res$df, nperm = 100L)
  #vimp1000_crps <- varimp(res$df, nperm = 1000L)

  save(vimp_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp_crps.rda")
  #save(vimp10_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp10_crps.rda")
  #save(vimp100_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp100_crps.rda")
  #save(vimp1000_crps, file = "~/svn/partykit/pkg/disttree/inst/draft/vimp1000_crps.rda")

@@

@@

\begin{figure}[t!]
%\begin{subfigure}{0.5\textwidth}
<<rain_axams_varim_margeffects1_mu, fig=TRUE, echo=FALSE>>=
par(mar = c(2,10,2,2))
barplot(sort(vimp_crps, decreasing = TRUE)[1:10], horiz = TRUE, las = 1, axes = FALSE)
@
%\end{subfigure}
%\hspace{-1cm}
%\begin{subfigure}{0.5\textwidth}
%<<rain_axams_varim_margeffects1_sigma, fig=TRUE, echo=FALSE>>=
%plot(x = sp1[,id_mi[1]], y = sp1[, "sigma"], type = "l", xlab = names_mi[1], ylab = %"sigma")
%@
%\end{subfigure}
\caption{\label{varimp} Variable imortance: top 10 list for the forest model}
\end{figure}
